---
title: "p8122_HW1"
author: "yc4384_Yangyang_Chen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary library
library(dplyr)
library(knitr)
```

## Question 1a
```{r}
individuals = data.frame(
  Individual = 1:8,
  Y0 = c(0, 1, 0, 1, 1, 0, 1, 0), 
  Y1 = c(0, 0, 1, 0, 0, 1, 0, 0)
)

# calculate treatment effects
individuals = individuals |>
  mutate(TE = Y1 - Y0) # add a col TE

individuals |>
  kable()
```

## Question 1b
```{r}
ACE = individuals |>
  summarise(ACE = mean(TE)) |>
  pull(ACE)

cat("The average causal effect (ACE) is:", ACE, "\n")
```

## Question 1c
```{r}
assigned_treatment = c(1, 0, 1, 1, 0, 0, 0, 1)

individuals = individuals |>
  mutate(Assigned_Treatment = assigned_treatment,
         Observed = ifelse(assigned_treatment == 1, Y1, Y0))

cat("Table: Observed Outcomes under Assigned Treatment\n")
individuals |>
  select(Individual, Y0, Y1, Assigned_Treatment, Observed) |>
  kable()

mean_treatment = individuals |>
  filter(Assigned_Treatment == 1) |>
  summarise(mean_treatment = mean(Observed)) |>
  pull(mean_treatment)

mean_control = individuals |>
  filter(Assigned_Treatment == 0) |>
  summarise(mean_control = mean(Observed)) |>
  pull(mean_control)

association = mean_treatment - mean_control
cat("The association between treatment and outcome under specific treatment assignment is:", association, "\n")
```
Interpretation:

The association between treatment and outcome under the specific treatment assignment is `r association`, which indicates individuals in the treatment group had worse health status (mean = 0.25) compared to the control group (mean = 0.75).

## Question 1d
```{r}
set.seed(329) 
random_assignment = sample(c(0, 1), size = 8, replace = TRUE)

individuals = individuals |>
  mutate(Random_Assigned_Treatment = random_assignment,
         Random_Assigned_Observed = ifelse(random_assignment == 1, Y1, Y0))

cat("Table: Observed Outcomes under Random Assignment\n")
individuals |>
  select(Individual, Y0, Y1, Random_Assigned_Treatment, Random_Assigned_Observed) |>
  kable()

# Calculate the association for random assignment
mean_treatment_random = individuals |>
  filter(Random_Assigned_Treatment == 1) |>
  summarise(mean_treatment_random = mean(Random_Assigned_Observed)) |>
  pull(mean_treatment_random)

mean_control_random = individuals |>
  filter(Random_Assigned_Treatment == 0) |>
  summarise(mean_control_random = mean(Random_Assigned_Observed)) |>
  pull(mean_control_random)

association_random = mean_treatment_random - mean_control_random
cat("The association between treatment and outcome under random assignment is:", association_random, "\n")

# Compare with part 1b result (ACE)
cat("The ACE calculated in part 1b is:", ACE, "\n")
cat("Comparison: Association under random assignment:", association_random, "vs ACE:", ACE, "\n")
```

## Question 2a

The unit is the patient who is being treated for blood pressure management.

## Question 2b

The treatment is the dosage of medication: 'High Dose' or 'Low Dose'

## Question 2c

In this case, we could define two potential outcomes:

*	 $Y_{\text{high}}$ : Blood pressure could be too high if the patient continues with the high dose of medication.
*	 $Y_{\text{low}}$ : Blood pressure could be perfect (or controlled) under the low dose of medication.

## Question 2d
```{r}
outcomes_numeric = data.frame(
  Treatment = c("High Dose", "Low Dose"),
  Y_high = c(0, NA),   
  Y_low = c(NA, 1)      
)

# Calculate causal effect: Y_low - Y_high
causal_effect = outcomes_numeric$Y_low[2] - outcomes_numeric$Y_high[1]
causal_effect
```

Interpretation: Switching to the low dose results in better control of blood pressure.


## Question 2e

SUTVA is plausible in this scenario:
* No interference: The patient's outcome is not influenced by others' treatment.
* Consistency: The low dose consistently results in better blood pressure control.

## Question 2f

* SUTVA must be plausible to ensure that the potential outcome framework is valid.
* This ensures that the treatment's effect is not influenced by other external factors and the outcome is consistent across treatments.

# Question 2g

* Probabilistic: No, the assignment is deterministic based on blood pressure readings.
* Individualistic: Yes, the treatment is specifically tailored for this patient.
* Unconfounded: Yes, treatment is based directly on the observed blood pressure levels, minimizing confounding.
* Controlled: Yes, the physician adjusts the dosage in a controlled manner based on the outcome.

## Question 2h
```{r}
set.seed(123) 

random_assignment = sample(c("High Dose", "Low Dose"), size = 1, replace = TRUE)
cat("Random Assignment: The patient is assigned to:", random_assignment, "\n")

random_outcome = ifelse(random_assignment == "High Dose", "Too High", "Perfect")
cat("Observed outcome under random assignment:", random_outcome, "\n")

n = 10  # Number of random assignments
results = data.frame(Trial = 1:n, Assigned_Treatment = NA, Outcome = NA)

for (i in 1:n) {
  assigned_treatment = sample(c("High Dose", "Low Dose"), size = 1, replace = TRUE)
  observed_outcome = ifelse(assigned_treatment == "High Dose", "Too High", "Perfect")
  results$Assigned_Treatment[i] = assigned_treatment
  results$Outcome[i] = observed_outcome
}

cat("\nTable: Outcomes under repeated random assignment\n")
print(results)
```


